---
title: "Intervention in augmented reality"
subtitle: "Analysis script for a preregistration"
author: "Removed for anonymization"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(message=FALSE,
                      warning=FALSE,
                      cache.lazy=FALSE)
```

# Dependencies

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

# get data
```{r}
data_processed <- read_csv("../../data/processed/data_processed.csv")
```

# Demographics
```{r}
gender <- data_processed |>
  count(gender)

mean_age <- mean(data_processed$age, na.rm = TRUE)


bedingung <- data_processed |>
  count(bedingung)

```

# Statistics
#### Evaluate H1 vs its complement with GORIC: ANOVA Example ####

# Load libraries. 
# These contain functions such as 'goric' that will be used in this R code. 
# Each time you re-open this R file you have to execute this step.

## First, install the packages, if you have not done this already:
```{r}
if (!require("psych")) install.packages("psych")
if (!require("restriktor")) install.packages("restriktor")
```

## Then, load the packages:
```{r}
library(psych) # for the function describeBy
library(restriktor) # for the goric function
```

# If you want to use restriktor from github:
#if (!require("devtools")) install.packages("devtools")
#library(devtools) 
#install_github("LeonardV/restriktor")
#library(restriktor) # for goric function

################################################################################


# Notably, it is only possible to load the data if you are using the correct 
# working directory (with both your R script and data file). 
# The command `getwd()` shows you your current working directory.
#
# You can change the working directory to the one you prefer using the function 
# `setwd()` by specifying the correct location between parentheses. 
# Alternatively, in Rstudio, you can use the "Session" tab (on top) or you can 
# use the "Files"-pane (on top of probably the right lower box of Rstudio, this 
# pane is located next to the panes for "Plots", "Packages", "Help", and "Viewer").


```{r}

```

